---
title: "matrixPriorSimNotebook"
format: 
  html:
    theme: Sandstone
    toc: true
    code-fold: true
    code-tools: true
    code-block-border-left: true
    code-block-bg: true
    self-contained: true
editor: source
---

# Load



```{r}
#| warning: false

version
library(ggplot2)
library(latex2exp)
ggplot2::set_theme(theme_bw())
library(PMA)
library(softImpute)

source('./matrixPrior_utils.R') # util functions for running the simulation


save_folder = "C:/Users/Cathe/Documents/School/Genetic Hypothesis Testing using Negative Controls/multilevel-treatment-multivariable-outcome-github/multilevel-treatment-multivariable-outcome/plots/matrixPrior/sim/"
```


# Setup

True `effects' matrix $M \in \mathbb{R}^{n\times m}$ with rank $r\in\mathbb{N}$.

Observed values $Y=M+E$ with independent noise $E_{ij} \sim N(0, \sigma_{ij}^2)$.





# Simulation with Estimating Structured Matrix

```{r}
#| label: testMatrixEstimation
#| output: false
#| warning: false

################################## Test Matrix Estimation ################################################# 
# test methods of estimating the structured matrix
# SVD, sparse SVD (Mat Decomp)
# linear reg (our implementation), and softImpute (Mat Comp where there is 1 element missing) 

{
r = 5
n = 15
m = 10
set.seed(12345)
# M = create_lowrank_matrix(r=r, n=n, m=m)
M = create_blocky_matrix(r=r, n=n, m=m) 
svd_maxrank = r
# noise_sd_s = seq(from = 0, to = .5, length.out = 5)
noise_sd_s = seq(from = 0, to = 2, length.out = 5)

M_obs_plots = list()
noise_label_plots = list()
X_mcomp_linearreg_list = list()
X_mcomp_linearreg_plots = list()
X_mcomp_softImpute_list = list()
X_mcomp_softImpute_plots = list()
X_svd_plots = list()
X_ssvd_plots = list()

for(noise_sd_s_ in noise_sd_s) {
  print(noise_sd_s_)
  # observed Matrices
  noise_sd = rep(noise_sd_s_, times = n*m) 
  M_obs    = M + matrix(rnorm(n = length(noise_sd), mean = 0, sd = noise_sd),
                     nrow = n, ncol = m, byrow = TRUE) # display_matrix(M_obs) 
  noise_label_plots[[as.character(noise_sd_s_)]] <- local({
    noise_sd_s_ <- noise_sd_s_
    ggplot() + geom_text(aes(x=0, y=0, label = sprintf('noise sd:\n%.2f', noise_sd_s_)), size = 15) + theme_void()
  })
  # noise_sd_s_text = sprintf('noise sd:\n%.2f', noise_sd_s_)
  # noise_label_plots[[as.character(noise_sd_s_)]] = ggplot() + geom_text(aes(x=0, y=0, label = noise_sd_s_text), size = 15) + theme_void()
  
  
  
  # === Construct X matrices from matrix completion (linear reg, softImpute)
  # --- Linear Reg
  X_mcomp_linearreg_list[[as.character(noise_sd_s_)]] = matrix(NA, nrow = n, ncol = m)
  for(i in 1:n) {
    for(j in 1:m) {
      # LR: X is linearreg
      X_mcomp_linearreg_list[[as.character(noise_sd_s_)]][i,j] = linearreg(M_obs, i, j)
    }
  }
  M_obs_plots[[as.character(noise_sd_s_)]] = display_matrix(M_obs)
  X_mcomp_linearreg_plots[[as.character(noise_sd_s_)]] = display_matrix(X_mcomp_linearreg_list[[as.character(noise_sd_s_)]])
  
  
  # --- softImpute
  X_mcomp_softImpute_list[[as.character(noise_sd_s_)]] = matrix(NA, nrow = n, ncol = m)
  matcomp_fun = create_matrix_completion_softImpute(X = M_obs, rank_max = svd_maxrank, type = 'als')
  for(i in 1:n) {
    for(j in 1:m) {
      # LR: X is softImpute(d)
      X_mcomp_softImpute_list[[as.character(noise_sd_s_)]][i,j] = matcomp_fun(i, j)
    }
  }
  M_obs_plots[[as.character(noise_sd_s_)]] = display_matrix(M_obs)
  X_mcomp_softImpute_plots[[as.character(noise_sd_s_)]] = display_matrix(X_mcomp_softImpute_list[[as.character(noise_sd_s_)]])
  
  # === Construct X matrices from decomp (svd, sparse svd, etc...)
  # --- SVD
  svdres = svd(M_obs, nu = svd_maxrank, nv = svd_maxrank)
  X_svd  = svdres$u %*% diag(svdres$d[1:svd_maxrank]) %*% t(svdres$v)
  X_svd_plots[[as.character(noise_sd_s_)]] = display_matrix(X_svd)
  
  # --- Sparse SVD
  cv.out = PMA::SPC.cv(M_obs, sumabsvs = seq(1.2, min(5, sqrt(n), sqrt(m)), len = 10))
  pmd_res = PMA::SPC(M_obs, sumabsv=cv.out$bestsumabsv, K = svd_maxrank)
  X_sparse_svd = pmd_res$u %*% diag(pmd_res$d) %*% t(pmd_res$v)
  X_ssvd_plots[[as.character(noise_sd_s_)]] = display_matrix(X_sparse_svd)
  
  rm(noise_sd_s_)
}



# plot labels
p_label_Mobs  = ggplot() + geom_text(aes(x=0, y=0, label = 'Observed\nMatrix'), size = 12) + theme_void()
p_label_Mcomp_linearreg  = ggplot() + geom_text(aes(x=0, y=0, label = 'Matrix Completion\nLinear Reg'), size = 12) + theme_void()
p_label_Mcomp_softImpute = ggplot() + geom_text(aes(x=0, y=0, label = 'Matrix Completion\nsoftImpute'), size = 12) + theme_void()
p_label_SVD = ggplot() + geom_text(aes(x=0, y=0, label = 'Matrix Decomp\nSVD'), size = 12) + theme_void()
p_label_SSVD = ggplot() + geom_text(aes(x=0, y=0, label = 'Matrix Decomp\nSparse SVD'), size = 12) + theme_void()


layout_matrix = cbind( matrix(c(NA, 6*length(noise_sd_s)+(1:5)),
                              nrow = 6),
                       matrix(1:(6*length(noise_sd_s)), 
                              nrow = 6, byrow = TRUE))

grob <- gridExtra::arrangeGrob(grobs = c(noise_label_plots, 
                                         M_obs_plots, 
                                         X_mcomp_linearreg_plots, 
                                         X_mcomp_softImpute_plots,
                                         X_svd_plots,
                                         X_ssvd_plots,
                                         p_label_Mobs, p_label_Mcomp_linearreg, p_label_Mcomp_softImpute, p_label_SVD, p_label_SSVD),
                               layout_matrix = layout_matrix)



gridExtra::grid.arrange(grob)
ggsave(sprintf('%s/test_matcomp_linearreg.pdf', save_folder), grob, width = 24, height = 16)

}
```


```{r}
#| label: fig-testMatrixEstimation
#| fig-cap: "Check Matrix Estimation Across various Methods and Noise Levels"
#| fig.height: 16
#| fig.width: 24
#| echo: false

gridExtra::grid.arrange(grob)
```



# Simulation with Hypothesis Testing

## 1 Run

### Setting A: 

```{r}
#| label: matrixpriorsimA
#| output: false
#| warning: false

{
# A: Easy case- large signal, small noise
r = 5
n = 15
m = 10
set.seed(12345)
M        = create_blocky_matrix(r=r, n=n, m=m)     # the signal
noise_sd = runif(n*m, min=.1, max=.5)              # sd of the noise


grobA = sim_and_plot_samplesplit(M=M, noise_sd=noise_sd, svd_maxrank=r,
                                 save_folder=sprintf('%s/A/', save_folder))

grobA_nosamplesplit = sim_and_plot_nosamplesplit(M=M, noise_sd=noise_sd/sqrt(2), svd_maxrank=r,
                                                 save_folder=sprintf('%s/A/', save_folder))
}
```



```{r}
#| label: fig-A-split
#| fig-cap: "Setting A with Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobA)
```

```{r}
#| label: fig-A-nosplit
#| fig-cap: "Setting A without Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobA_nosamplesplit)
```



### Setting B: 

```{r}
#| label: matrixpriorsimB
#| output: false
#| warning: false

{
# B: Medium case- ----- signal, medium noise
r = 5
n = 15
m = 10
set.seed(12345)
M        = create_blocky_matrix(r=r, n=n, m=m)    # the signal
noise_sd = runif(n*m, min=.2, max=2)              # sd of the noise


grobB = sim_and_plot_samplesplit(M=M, noise_sd=noise_sd, svd_maxrank=r,
                     save_folder=sprintf('%s/B/', save_folder))

grobB_nosamplesplit = sim_and_plot_nosamplesplit(M=M, noise_sd=noise_sd/sqrt(2), svd_maxrank=r,
                                                 save_folder=sprintf('%s/B/', save_folder))
}
```



```{r}
#| label: fig-B-split
#| fig-cap: "Setting B with Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobB)
```

```{r}
#| label: fig-B-nosplit
#| fig-cap: "Setting B without Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobB_nosamplesplit)
```



### Setting C: 

```{r}
#| label: matrixpriorsimC
#| output: false
#| warning: false

{
# C: Hard case- ----- signal, large noise
r = 5
n = 15
m = 10
set.seed(12345)
M        = create_blocky_matrix(r=r, n=n, m=m)  # the signal
noise_sd = runif(n*m, min=3, max=8)            # sd of the noise


grobC = sim_and_plot_samplesplit(M=M, noise_sd=noise_sd, svd_maxrank=r,
                     save_folder=sprintf('%s/C/', save_folder))

grobC_nosamplesplit = sim_and_plot_nosamplesplit(M=M, noise_sd=noise_sd/sqrt(2), svd_maxrank=r,
                                                 save_folder=sprintf('%s/C/', save_folder))

}
```



```{r}
#| label: fig-C-split
#| fig-cap: "Setting C with Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobC)
```

```{r}
#| label: fig-C-nosplit
#| fig-cap: "Setting C without Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobC_nosamplesplit)
```



### Setting D: 

```{r}
#| label: matrixpriorsimD
#| output: false
#| warning: false

{
# D: Larger matrix case- strong signal, small noise, large matrix
r = 5
n = 100
m = 50
set.seed(12345)
M        = create_blocky_matrix(r=r, n=n, m=m)*5  # the signal
noise_sd = runif(n*m, min=.01, max=5)             # sd of the noise


grobD = sim_and_plot_samplesplit(M=M, noise_sd=noise_sd, svd_maxrank=r,
                     save_folder=sprintf('%s/D/', save_folder))

grobD_nosamplesplit = sim_and_plot_nosamplesplit(M=M, noise_sd=noise_sd/sqrt(2), svd_maxrank=r,
                                                 save_folder=sprintf('%s/D/', save_folder))
}
```



```{r}
#| label: fig-D-split
#| fig-cap: "Setting D with Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobD)
```

```{r}
#| label: fig-D-nosplit
#| fig-cap: "Setting D without Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobD_nosamplesplit)
```


### Setting E: 

```{r}
#| label: matrixpriorsimE
#| output: false
#| warning: false

{
# E: harder Larger matrix case- strong signal, medium noise, large matrix
r = 5
n = 100
m = 50
set.seed(12345)
M        = create_blocky_matrix(r=r, n=n, m=m)*5  # the signal
noise_sd = runif(n*m, min=1, max=8)            # sd of the noise

grobE = sim_and_plot_samplesplit(M=M, noise_sd=noise_sd, svd_maxrank=r,
                     save_folder=sprintf('%s/E/', save_folder))

grobE_nosamplesplit = sim_and_plot_nosamplesplit(M=M, noise_sd=noise_sd/sqrt(2), svd_maxrank=r,
                                                 save_folder=sprintf('%s/E/', save_folder))
}
```



```{r}
#| label: fig-E-split
#| fig-cap: "Setting E with Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobE)
```

```{r}
#| label: fig-E-nosplit
#| fig-cap: "Setting E without Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobE_nosamplesplit)
```




### Setting F: 

```{r}
#| label: matrixpriorsimF
#| output: false
#| warning: false

{
# F: hardest Larger matrix case- strong signal, very large noise, large matrix
r = 5
n = 100
m = 50
set.seed(12345)
M        = create_blocky_matrix(r=r, n=n, m=m)*5  # the signal
noise_sd = runif(n*m, min=3, max=12)            # sd of the noise


grobF = sim_and_plot_samplesplit(M=M, noise_sd=noise_sd, svd_maxrank=r,
                                 save_folder=sprintf('%s/F/', save_folder))

grobF_nosamplesplit = sim_and_plot_nosamplesplit(M=M, noise_sd=noise_sd/sqrt(2), svd_maxrank=r,
                                                 save_folder=sprintf('%s/F/', save_folder))
}
```



```{r}
#| label: fig-F-split
#| fig-cap: "Setting F with Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobF)
```

```{r}
#| label: fig-F-nosplit
#| fig-cap: "Setting F without Sample Splitting"
#| fig.width: 15
#| fig.height: 12 
#| echo: false
gridExtra::grid.arrange(grobF_nosamplesplit)
```






## Many Runs


How do we `average' across multiple runs?

eg. suppose we have a vector of pvals from multiple runs: pval1, pval2, ..., pvalK. 


```{r}
rep_times = 20
```


### Setting A

```{r}
#| output: false
#| warning: false


# A: Easy case- large signal, small noise
r = 5
n = 15
m = 10
set.seed(12345)
M        = create_blocky_matrix(r=r, n=n, m=m)     # the signal
noise_sd = runif(n*m, min=.1, max=.5)              # sd of the noise


pval_list_samplesplit   = list()
pval_list_nosamplesplit = list()
for(k in 1:rep_times) {
  grobA = sim_and_plot_samplesplit(M=M, noise_sd=noise_sd, svd_maxrank=r,
                                 save_folder=sprintf('%s/A/', save_folder), 
                                 return_pvals = TRUE)

  grobA_nosamplesplit = sim_and_plot_nosamplesplit(M=M, noise_sd=noise_sd/sqrt(2), svd_maxrank=r,
                                                   save_folder=sprintf('%s/A/', save_folder), 
                                                   return_pvals=TRUE)
  pval_list_samplesplit[[k]]   = grobA$pvals
  pval_list_nosamplesplit[[k]] = grobA_nosamplesplit$pvals
}



# just take the average????
avg_pval_list_samplesplit   = list()

for(method in names(pval_list_samplesplit[[1]])) {
  matrix_sum_samplesplit   = matrix(0, nrow = n, ncol = m)

  for(k in 1:rep_times) {
    matrix_sum_samplesplit    =   matrix_sum_samplesplit   + pval_list_samplesplit[[k]][[method]]
    
  }
  avg_pval_list_samplesplit[[method]]   =   matrix_sum_samplesplit / rep_times
}

avg_pval_list_nosamplesplit = list()
for(method in names(pval_list_nosamplesplit[[1]])) {
  
  matrix_sum_nosamplesplit = matrix(0, nrow = n, ncol = m)
  for(k in 1:rep_times) {
   
    matrix_sum_nosamplesplit  = matrix_sum_nosamplesplit + pval_list_nosamplesplit[[k]][[method]]
  }
  avg_pval_list_nosamplesplit[[method]] = matrix_sum_nosamplesplit / rep_times
}


# taking a mean of uniform(0,1) is not uniform (ie even if we had a valid pvalue, taking the mean over many runs is not ideal)
take_mean_of_unif <- function() {mean(runif(n=10, min = 0, max = 1))} # eg repeat sim X times, then take the avg pvals (should be close to .5, converges to .5 as nreps --> inft by lln)

qqplot(replicate(100, expr = take_mean_of_unif()), runif(10000, min=0, max=1)); abline(a=0, b=1)
qqplot(x = replicate(100, expr = take_mean_of_unif()), distribution = qunif)
qqplot(qunif(ppoints(500), min=0, max=1),  y = replicate(100, expr = take_mean_of_unif())); abline(a=0, b=1)
hist(replicate(100, expr = take_mean_of_unif()), xlim = c(0, 1))

# maybe take the median? This is also not good, this would also --> .5 as nreps --> infty
take_median_of_unif <- function() {median(runif(n=10, min = 0, max = 1))} # eg repeat sim X times, then take the avg pvals (should be close to .5, converges to .5 as nreps --> inft by lln)
hist(replicate(100, expr = take_median_of_unif()), xlim = c(0, 1))


# maybe just combine them all together?


# put into array format
pval_array_samplesplit = array(data=NA, dim = c(n, m, rep_times))
pval_array_samplesplit[, , 2] = pval_list_nosamplesplit[[k]][[method]]


plot_pvals_many(P = avg_pval_list_samplesplit,
                M = M,  
                method_colors = c('mediumpurple3', 'dodgerblue2', 'dodgerblue4', 'palegreen2', 'palegreen4', 'orangered2'))

  
plot_pvals_many(P = avg_pval_list_nosamplesplit,
                M = M,  
                method_colors = c('mediumpurple3', 'dodgerblue2', 'dodgerblue4', 'palegreen2', 'palegreen4', 'orangered2'))

```


```{r}

Ps = pval_list_samplesplit

#' @param Ps list of matrices of pvals to plot, named
#' @param M true params (need to know which are actually 0) or matrix where 0s are in the right pos
#' @param method_colors (vector) of color names for each of the method categories
#' @param Ps_has_many_reps (boolean) whether Ps (list of pvalue matrices) has many repetitions done
#'        if  TRUE, Ps is a list of list of matrices
#'        if FALSE, Ps is a list of matrices
plot_pvals_many <- function(Ps, M, method_colors=NULL, Ps_has_many_reps=FALSE) {
  # P = pvals_svd
  # P = pvals_standard
  
  
  df = NULL
  if(Ps_has_many_reps) {
    for(k in 1:length(Ps)) {
        for(p_name in names(Ps[[k]])) {
            P = Ps[[k]][[p_name]]
            dfP = reshape2::melt(P, c("x", "y"), value.name = "pval")
            dfM = reshape2::melt(M, c("x", "y"), value.name = "theta")
            df_ = merge(dfP, dfM, by = c('x', 'y'))
            df_$method = p_name
            df = rbind(df, df_)
        }
    }
    df$method = factor(df$method, levels = names(P))
  } else {
    for(p_name in names(Ps)) {
      P = Ps[[p_name]]
      dfP = reshape2::melt(P, c("x", "y"), value.name = "pval")
      dfM = reshape2::melt(M, c("x", "y"), value.name = "theta")
      df_ = merge(dfP, dfM, by = c('x', 'y'))
      df_$method = p_name
      df = rbind(df, df_)
      
    }
    df$method = factor(df$method, levels = names(Ps))
  }
  
  
  
  
  

  
  # null (theta = 0)
  p_null = ggplot(df |> dplyr::filter(theta == 0) |> 
                    dplyr::group_by(method) |>
                    dplyr::arrange(pval) |> 
                    dplyr::mutate(rej_rate = (1:dplyr::n())/dplyr::n()),
                  aes(x = pval, y = rej_rate, group = method, color = method)) + 
    geom_abline(aes(slope = 1, intercept = 0)) +
    geom_point(alpha = .8, size = .6, key_glyph = "rect") +
    geom_line() +
    # scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
    scale_color_discrete(palette = method_colors) + # NULL uses default colors
    labs(title = 'Rejection Rate by p-value (qqplot)',
         subtitle = '(when Theta=0)',
         x = 'p-value', y = 'Rejection Rate')
  
  # alt (theta != 0)
  p_alt = ggplot(df |> dplyr::filter(theta != 0) |> 
                   dplyr::group_by(method) |>
                   dplyr::arrange(pval) |> 
                   dplyr::mutate(rej_rate = (1:dplyr::n())/dplyr::n()),
                 aes(x = pval, y = rej_rate, group = method, color = method)) + 
    geom_abline(aes(slope = 1, intercept = 0)) +
    geom_point(alpha = .8, size = .6, key_glyph = "rect") +
    geom_line() +
    # scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
    scale_color_discrete(palette = method_colors) + # NULL uses default colors
    labs(title = 'Rejection Rate by p-value (qqplot)',
         subtitle = '(when Theta!=0)',
         x = 'p-value', y = 'Rejection Rate')
  
  
  return(list(p_null=p_null, p_alt=p_alt))
}

  
```

















